---
title: "Tables in 'Accuracy Is Not All You Need', Extracting Fairness Metrics (TP/FN/FP/PRECISION/RECALL/F1_SCORE)"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

This notebook contains code for producing the table in the paper. 

```{r}
library(tidyverse)
library(kableExtra)
library(janitor)
```

# Define Functions 
## Define Scoring Function (Pairwise Bonferroni T-tests)
```{r scoring summary functions}
# t-test function for the score_to_df function
score_t_test = function(mdl, augmenter, default, data, score){
  d = data %>% filter(model == mdl)

  x = d[[score]][d$augmenter == augmenter]
  mu = d[[score]][d$augmenter == default]
  
  if (length(mu) == 1){
    t = t.test(x = x,
           mu = mu, paired = FALSE, var.equal = FALSE,
           conf.level = 0.95)
  } else {
    t = t.test(x = x,
           y = mu, paired = FALSE, var.equal = FALSE,
           conf.level = 0.95) 
  }
  return (t)
}

# score_to_df function: bonferroni adjusted t-tests for all augmentations for all models against a baseline augmentation
score_to_df = function(data, score, baseline){
  dfs = NULL
  i = 1
  for(mdl in unique(data$model)){
    for (aug in unique(data$augmenter)){
      v = data[[score]][data$augmenter == aug & data$model == mdl]
      
      if (length(v) <= 2){
          dfs[[i]] = tibble(model=mdl, augmenter=aug, mean=v[1], sd=NA, 
                    conf_int = "",
                    p_value=NA)
          i = i+1
          next
      }
      
      mu = mean(v)
      sigma = sd(v)
      print(paste(mdl, aug, length(v)))
      t = score_t_test(mdl=mdl, augmenter=aug, default=baseline, data=data, score=score)
      
      p = p.adjust(t$p.value, method = "bonferroni", n = 6)

      dfs[[i]] = tibble(model=mdl, augmenter=aug, mean=mu, sd=sigma, 
                        conf_int = paste("(",  round(t$conf.int[1], 2),", ",  round(t$conf.int[2], 2), ")", sep = ""),
                        p_value=p)
      i = i+1
    }
  }
  return(bind_rows(dfs))
}
```

## Define function for calculating and creating table for chosen value 
```{r}
create_metric_table = function(metric_col_name, data, metric_label, order_groups, order_models, model_names_from_to, aug_names_from_to){
    #' Create metric table with selected metric for models 
    #'
    #' Create metric table with selected metric within a data table.Computes t-tests as defined in the scoring function "score_to_df"
    #' @param metric_col_name column in data -> the selected metric
    #' @param data dataframe -> the data with the metric
    #' @param order_groups list -> how the augmentation groups should be ordered
    #' @param order_models list -> how the models should be ordered
    #' @param model_names_from_to list -> containing two lists: change model names from (list) to model names (list)
    #' @param aug_names_from_to list -> containing two lists: change augmenter names from (list) to augmenter names (list)
    #' This part goes in the Details!
  
  # create metric df 
  metric_df <- data %>% 
  select(model, 
         metric_col_name,
         augmenter = augmenter
  ) %>% 
  filter(augmenter != "No augmentation") # rm single run with no augmentation

  # calculate scores 
  metric_df_scores <- score_to_df(metric_df, metric_col_name, "Danish names")
  
  # create performance table with p-values, standard deviation in parenthesis
  metric_table = metric_df_scores %>% arrange(factor(model, levels = order_models)) %>% 
    mutate(mean = paste(format(round_half_up(mean, 1), digits = 1, nsmall=1)), # add mean scores
           p_value_star = if_else(p_value < 0.05, "*", "", missing =""), # define star if p-val is significant
           string_value = if_else(is.na(sd), paste(mean, sep=""), # formatting sd and p-val
                                  paste(mean,  " (", 
                                        format(round_half_up(sd, 1), digits = 1, nsmall=1), 
                                        ")", p_value_star, sep="")), 
           model = plyr::mapvalues(model, from=model_names_from_to[[1]], to=model_names_from_to[[2]]),
           augmenter = plyr::mapvalues(augmenter, from=aug_names_from_to[[1]], to=aug_names_from_to[[2]])) %>% 
    select(-c(mean, sd, p_value, p_value_star, conf_int)) %>% 
    pivot_wider(names_from = augmenter, values_from=c(string_value)) %>% 
    select(Model=model, all_of(aug_names_from_to[[2]])) # reorder
  
  # add metric col, specifying the metric 
  Metric <- rep(metric_label, 10)
  metric_table <- cbind(metric_table, Metric)
  
  # re order columns 
  metric_table <- metric_table[, c("Model", "Metric", order_groups)]

  return(metric_table)
}
```

## Define Function For Calculating and Creating Tables for Multiple Metrics
```{r}
create_multiple_metric_tables <- function(metric_col_names, metric_labels, data, order_groups, order_models, model_names_from_to, aug_names_from_to) {
  # create empty list to store metric tables
  metric_tables <- list()
  
  # loop through each metric_col_name and metric_label
  for (i in seq_along(metric_col_names)) {
    # call create_metric_table on the current metric_col_name and metric_label
    current_metric_table <- create_metric_table(metric_col_names[[i]], data, metric_labels[[i]], order_groups, order_models, model_names_from_to, aug_names_from_to)
    
    # add the current metric table to the list
    metric_tables[[i]] <- current_metric_table
  }
  
  # combine all metric tables in the list into a single dataframe
  combined_metric_table <- do.call(rbind, metric_tables)
  
  # return the combined metric table
  return(combined_metric_table)
}

```

## Define function for collapsing model names (for table), so that it does not say model name for each metric 
```{r}
# collapse rows
collapse_rows_df <- function(df, variable){
  group_var <- enquo(variable)
  df %>%
    group_by(!! group_var) %>%
    mutate(groupRow = 1:n()) %>%
    ungroup() %>%
    mutate(!!quo_name(group_var) := ifelse(groupRow == 1, as.character(!! group_var), "")) %>%
    select(-c(groupRow))
}
```
Function credits: https://stackoverflow.com/questions/51450402/how-to-collapse-groups-of-row-values-within-a-table-using-formattable 

## All tables ! 
## Rename model and augmenters
```{r rename model and aug}
# original model names
mdl_names_from = c("spacy_large", "spacy_medium", "dacy_large", "dacy_medium", "danlp", "spacy_small", "flair", "dacy_small", "polyglot", "scandi_ner")

# new model names
mdl_names_to = c("SpaCy large (3.4.0)", "SpaCy medium (3.4.0)", "DaCy large (0.1.0)", "DaCy medium (0.1.0)", "DaNLP BERT", "SpaCy small (3.4.0)", "Flair", "DaCy small (0.1.0)", "Polyglot", "ScandiNER")

# aug names original
aug_names_from = c("Danish names", "Muslim names", "Female names", "Male names", "Muslim female names","Muslim male names", "Unisex names") 

# aug names new 
aug_names_to = c("Majority all", "Minority all", "Majority Women", "Majority Men", "Minority Women", "Minority Men", "Unisex")
```

## Define order of models and groups
```{r}
order_models <- c("scandi_ner","dacy_large", "dacy_medium", "dacy_small", "danlp", "flair", "spacy_large", "spacy_medium", "spacy_small", "polyglot")

order_groups = c("Majority all", "Minority all", "Majority Men", "Minority Men", "Majority Women", "Minority Women", "Unisex")
```

## Table 1: ALL ENTS EXCL MISC 
### Read in Files ! 
```{r}
ALL_EXCL_MISC_ents <- readbulk::read_bulk(directory = "results/ALL_EXCL_MISC", extension = ".csv")

ALL_EXCL_MISC_ents = ALL_EXCL_MISC_ents %>% 
  mutate(across(precision, ~ .x * 100), # mutate to go from vals such as 0.9 to 90 
         across(recall, ~ .x * 100), 
         across(F1_score, ~ .x * 100)
         )
```

### Table 1A: TP/FN/FP
#### Create metric_table: Calculate aggregates and do bonferroni testing
```{r}
#### CREATE TABLE #### 
metric_cols <- c("TP", "FN", "FP") # what the column names are called 
metric_labels <- c("TP", "FN", "FP") 

# create table 
ALL_EXCL_MISC_TABLE_1A <- create_multiple_metric_tables(metric_cols, # what the column names are called 
                                                     metric_labels, # what the metrics should be called in the table
                                                     ALL_EXCL_MISC_ents, # dataframe to take from 
                                                     order_groups, # order of groups (augmenters)
                                                     order_models, # order of models 
                                                     list(mdl_names_from, mdl_names_to), # how model names should be changed 
                                                     list(aug_names_from, aug_names_to) # how the augmenter names should be changed 
                                                     )

#### REORDER #### 

# define new order with the actual names 
new_order <- c("ScandiNER", "DaCy large (0.1.0)", "DaCy medium (0.1.0)", "DaCy small (0.1.0)", "DaNLP BERT", "Flair", "SpaCy large (3.4.0)", "SpaCy medium (3.4.0)", "SpaCy small (3.4.0)", "Polyglot")

# reorder so that precision and recall comes after each other 
ALL_EXCL_MISC_TABLE_1A <- ALL_EXCL_MISC_TABLE_1A[order(unlist(sapply(ALL_EXCL_MISC_TABLE_1A$Model, function(x) which(new_order == x)))),] 
```

#### Make Table (Visualisation)
```{r}
ALL_1A <- collapse_rows_df(ALL_EXCL_MISC_TABLE_1A, Model) %>% 
  kbl(
    booktabs=T, 
    col.names = c("Model", "Metric", "Majority", "Minority", "Majority", "Minority", "Majority", "Minority", "Majority"),
    caption = "<span style='font-size:18px'>Table 1A: NER PERFORMANCE (ALL EXCL MISC). MEAN TP/FN/FP. (Bonferroni corrected for each model for each metric)</span>",
    #format="latex",
    align=c("l", rep("c", nrow(.)-1)),
    table.attr = "style='width:90%;'"
    ) %>% 
  column_spec(column=1, bold=TRUE, border_right = T, width = "6em") %>% 
  column_spec(column=2, bold=TRUE, border_right = T, width = "4em") %>% 
  column_spec(3:9, width = "1em") %>%
  add_header_above(c(" " = 1, " " = 1, "All" = 2, "Men" = 2, "Women" = 2, "Unisex"=1)) %>% 
  kable_classic(full_width = T, html_font = "Cambria", font_size = 16, "striped", latex_options ="scale_down")

ALL_1A
```

### TABLE 1B: PRECISION, RECALL, F1
#### Create metric_table: Calculate aggregates and do bonferroni testing
```{r}
#### CREATE TABLE #### 
metric_cols <- c("precision", "recall", "F1_score") 
metric_labels <- c("Precision", "Recall", "F1") 

# create table 
ALL_EXCL_MISC_TABLE_1B <- create_multiple_metric_tables(metric_cols, # what the column names are called 
                                                     metric_labels, # what the metrics should be called in the table
                                                     ALL_EXCL_MISC_ents, # dataframe to take from 
                                                     order_groups, # order of groups (augmenters)
                                                     order_models, # order of models 
                                                     list(mdl_names_from, mdl_names_to), # how model names should be changed 
                                                     list(aug_names_from, aug_names_to) # how the augmenter names should be changed 
                                                     )

#### REORDER #### 

# define new order with the actual names 
new_order <- c("ScandiNER", "DaCy large (0.1.0)", "DaCy medium (0.1.0)", "DaCy small (0.1.0)", "DaNLP BERT", "Flair", "SpaCy large (3.4.0)", "SpaCy medium (3.4.0)", "SpaCy small (3.4.0)", "Polyglot")

# reorder so that precision and recall comes after each other 
ALL_EXCL_MISC_TABLE_1B <- ALL_EXCL_MISC_TABLE_1B[order(unlist(sapply(ALL_EXCL_MISC_TABLE_1B$Model, function(x) which(new_order == x)))),] 
```


#### Make Table (Visualisation)
```{r}
ALL_1B <- collapse_rows_df(ALL_EXCL_MISC_TABLE_1B, Model) %>% 
  kbl(
    booktabs=T, 
    col.names = c("Model", "Metric", "Majority", "Minority", "Majority", "Minority", "Majority", "Minority", "Majority"),
    caption = "<span style='font-size:18px'>Table 1B: NER PERFORMANCE (ALL EXCL MISC). MEAN PRECISION/RECALL/F1. (Bonferroni corrected for each model for each metric)</span>",
   #format="latex",
    align=c("l", rep("c", nrow(.)-1)),
    table.attr = "style='width:90%;'"
    ) %>% 
  column_spec(column=1, bold=TRUE, border_right = T, width = "6em") %>% 
  column_spec(column=2, bold=TRUE, border_right = T, width = "4em") %>% 
  column_spec(3:9, width = "1em") %>%
  add_header_above(c(" " = 1, " " = 1, "All" = 2, "Men" = 2, "Women" = 2, "Unisex"=1)) %>% 
  kable_classic(full_width = T, html_font = "Cambria", font_size = 16, "striped", latex_options ="scale_down")

ALL_1B
```

## Table 2: PER ENTS ONLY
### Read in File ! 
```{r read files}
PER_ents <- readbulk::read_bulk(directory = "results/PER", extension = ".csv")

PER_ents = PER_ents %>% 
  mutate(across(precision, ~ .x * 100),
         across(recall, ~ .x * 100),
         across(F1_score, ~ .x * 100)
         )
```

### TABLE 2A: 
#### Create metric_table: Calculate aggregates and do bonferroni testing
```{r}
#### CREATE TABLE #### 
metric_cols <- c("TP", "FN", "FP")
metric_labels <- c("TP", "FN", "FP")

PER_TABLE_2A <- create_multiple_metric_tables(metric_cols, 
                                           metric_labels, 
                                           PER_ents, 
                                           order_groups, 
                                           order_models, 
                                           list(mdl_names_from, mdl_names_to), 
                                           list(aug_names_from, aug_names_to))

#### REORDER ####

# define new order with the actual names 
new_order <- c("ScandiNER", "DaCy large (0.1.0)", "DaCy medium (0.1.0)", "DaCy small (0.1.0)", "DaNLP BERT", "Flair", "SpaCy large (3.4.0)", "SpaCy medium (3.4.0)", "SpaCy small (3.4.0)", "Polyglot")


# reorder so that precision and recall comes after each other 
PER_TABLE_2A <- PER_TABLE_2A[order(unlist(sapply(PER_TABLE_2A$Model, function(x) which(new_order == x)))),] 
```

#### Make Table
```{r}
PER_2A <- collapse_rows_df(PER_TABLE_2A, Model) %>% 
  kbl(
    booktabs=T, 
    col.names = c("Model", "Metric", "Majority", "Minority", "Majority", "Minority", "Majority", "Minority", "Majority"),
    caption = "<span style='font-size:18px'>Table 2A: PER ENT PERFORMANCE. MEAN TP/FN/FP. (Bonferroni corrected for each model for each metric)</span>",
    #format="latex",
    align=c("l", rep("c", nrow(.)-1)),
    table.attr = "style='width:90%;'"
    ) %>% 
  column_spec(column=1, bold=TRUE, border_right = T, width = "6em") %>% 
  column_spec(column=2, bold=TRUE, border_right = T, width = "4em") %>% 
  column_spec(3:9, width = "1em") %>%
  add_header_above(c(" " = 1, " " = 1, "All" = 2, "Men" = 2, "Women" = 2, "Unisex"=1)) %>% 
  kable_classic(full_width = T, html_font = "Cambria", font_size = 16, "striped", latex_options ="scale_down")

PER_2A
```


### TABLE 2B: TP/FN/FP
#### Create metric_table: Calculate aggregates and do bonferroni testing
```{r}
#### CREATE TABLE #### 
metric_cols <- c("precision", "recall", "F1_score") 
metric_labels <- c("Precision", "Recall", "F1") 

PER_TABLE_2B <- create_multiple_metric_tables(metric_cols, 
                                           metric_labels, 
                                           PER_ents, 
                                           order_groups, 
                                           order_models, 
                                           list(mdl_names_from, mdl_names_to), 
                                           list(aug_names_from, aug_names_to))

#### REORDER ####

# define new order with the actual names 
new_order <- c("ScandiNER", "DaCy large (0.1.0)", "DaCy medium (0.1.0)", "DaCy small (0.1.0)", "DaNLP BERT", "Flair", "SpaCy large (3.4.0)", "SpaCy medium (3.4.0)", "SpaCy small (3.4.0)", "Polyglot")


# reorder so that precision and recall comes after each other 
PER_TABLE_2B <- PER_TABLE_2B[order(unlist(sapply(PER_TABLE_2B$Model, function(x) which(new_order == x)))),] 
```

#### Make Table
```{r}
PER_2B <- collapse_rows_df(PER_TABLE_2B, Model) %>% 
  kbl(
    booktabs=T, 
    col.names = c("Model", "Metric", "Majority", "Minority", "Majority", "Minority", "Majority", "Minority", "Majority"),
    caption = "<span style='font-size:18px'>Table 2B: PER ENT PERFORMANCE. MEAN TP/FN/FP. (Bonferroni corrected for each model for each metric</span>)",
    #format="latex",
    align=c("l", rep("c", nrow(.)-1)),
    table.attr = "style='width:90%;'"
    ) %>% 
  column_spec(column=1, bold=TRUE, border_right = T, width = "6em") %>% 
  column_spec(column=2, bold=TRUE, border_right = T, width = "4em") %>% 
  column_spec(3:9, width = "1em") %>%
  add_header_above(c(" " = 1, " " = 1, "All" = 2, "Men" = 2, "Women" = 2, "Unisex"=1)) %>% 
  kable_classic(full_width = T, html_font = "Cambria", font_size = 16, "striped", latex_options ="scale_down")

PER_2B
```



### Save Tables to CSV
```{r}
write_csv(ALL_EXCL_MISC_TABLE_1A, "tables/1A_NER_PERFORMANCE.csv")
write_csv(ALL_EXCL_MISC_TABLE_1B, "1B_NER_PERFORMANCE.csv")

write_csv(PER_TABLE_2A, "2A_PER_ENTS.csv")
write_csv(PER_TABLE_2B, "2B_PER_ENTS.csv")
```


## Remove SD and Stars for raw vals 
```{r}
# define a function to remove parentheses and stars from a string
remove_parentheses_and_star <- function(x) {
  gsub("\\s*\\([^\\)]+\\)|\\*", "", x)
}
columns <- c("Majority all", "Minority all", "Majority Women", "Majority Men", "Minority Women", "Minority Men", "Unisex")

# apply the function to multiple columns using lapply()
ALL_EXCL_MISC_TABLE_1A[, columns] <- lapply(ALL_EXCL_MISC_TABLE_1A[, columns], remove_parentheses_and_star)

# print the updated dataframe
print(ALL_EXCL_MISC_TABLE_1A)
```

```{r}
write_csv(ALL_EXCL_MISC_TABLE_1A, "1A_NER_PERF_WITHOUT_SD.csv")
```


